\chapter{Publication List}
\label{ch:publication-list}

\section{Publications included in this Thesis}


\begin{itemize}
    \item \underline{C. Geckeler}, et al. "Autonomous Surface eDNA Collection from Rainforests: ETH BIODIVX in the XPrize Rainforest Finals", Under Review, 2024,
    \item \underline{C. Geckeler}, S. Heinrich, S. Mintchev, "Robotic Environmental Monitoring using Gelatin Hydrogels as a Biodegradable Adhesive", Under Review, 2024,
    \item M. Muglikar, N. Neugebauer, \underline{C. Geckeler}, S. Mintchev, D. Scaramuzza, "Event Spectroscopy: Event-based Multispectral and Depth Sensing using Structured Light", Under Review, 2024
    \item \cite{Kirchgeorg2024} S. Kirchgeorg, J. Chang, Y. Ip, M. Jucker, \underline{C. Geckeler}, M. Lüthi, E. van der Loo, E. Mächler, N. Franco-Sierra, M. Herrera, L. Pellissier, K. Deiner, A. Desiderato, S. Mintchev, “eProbe: Sampling of Environmental DNA within Tree Canopies with Drones,” Environ. Sci. Technol., Sep. 2024,
    \item \cite{Geckeler2024a} \underline{C. Geckeler}, I. Kong, and S. Mintchev, “User-Centric Payload Design and Usability Testing for Agricultural Sensor Placement and Retrieval using Off-the-Shelf Micro Aerial Vehicles,” in 2024 21st International Conference on Ubiquitous Robots (UR), Jun. 2024, pp. 184–191.
    \item \cite{Geckeler2024} \underline{C. Geckeler}, E. Aucone, Y. Schnider, A. Simeon, J. von Bassewitz, Y. Zhu, S. Mintchev, “Learning Occluded Branch Depth Maps in Forest Environments Using RGB-D Images,” IEEE Robot. Autom. Lett., vol. 9, no. 3, pp. 2439–2446, Mar. 2024,
    \item \cite{Geckeler2023a} \underline{C. Geckeler}, S. E. Ramos, M. C. Schuman, and S. Mintchev, “Robotic Volatile Sampling for Early Detection of Plant Stress: Precision Agriculture Beyond Visual Remote Sensing,” IEEE Robot. Autom. Mag., pp. 2–12, 2023,
    \item \cite{Geckeler2023b} \underline{C. Geckeler}, B. A. Pizzani, and S. Mintchev, “Biodegradable Origami Gripper Actuated with Gelatin Hydrogel for Aerial Sensor Attachment to Tree Branches,” in Proceedings - IEEE International Conference on Robotics and Automation, May 2023, vol. 2023-May, pp. 5324–5330.
    \item \cite{Geckeler2022a} \underline{C. Geckeler} and S. Mintchev, “Bistable Helical Origami Gripper for Sensor Placement on Branches,” Adv. Intell. Syst., vol. 4, no. 10, p. 2200087, Oct. 2022,    
\end{itemize}

\section{Other Publications}

\begin{itemize}
    \item \cite{Cuniato2023} E. Cuniato, \underline{C. Geckeler}, M. Brunner, D. Strübin, E. Bähler, F. Ospelt, M. Tognon, S. Mintchev, R. Siegwart, “Design and Control of a Micro Overactuated Aerial Robot with an Origami Delta Manipulator,” in 2023 IEEE International Conference on Robotics and Automation (ICRA), May 2023, vol. 2023-May, pp. 5352–5358.
    \item \cite{Aucone2024} E. Aucone, \underline{C. Geckeler}, D. Morra, L. Pallottino, and S. Mintchev, “Synergistic morphology and feedback control for traversal of unknown compliant obstacles with aerial robots,” Nat. Commun., vol. 15, no. 1, p. 2646, Mar. 2024,
\end{itemize}


\section{Conference and Competition Attendances}

% MISC awards (startup campus etc)

\section{Supervised Students}

\subsection{Master Thesis}
Master students, 6 months full time, 30 ECTS


% TODO Maybe add student university (ie ETH, university of pisa, etc)
\begin{itemize}
    \item Daniele Morra, Spring 2022, "Model Predictive Controller for Path Following during Physical Interaction with Flexible Unknown Environments", supervised by Emaunele Aucone and Christian Geckeler, part of publication \cite{Aucone2024}

    \item  Madeleine Schuler, Autumn 2022, "Path planning and obstacle avoidance for an aerial-tethered robot in cluttered environments", supervised by Steffen Kirchgeorg and Christian Geckeler.

    \item Niklas Neugebauer, Autumn 2022, "Event-based structured light for depth sensing and material differentiation in forest canopies", supervised by Christian Geckeler and Manasi Muglikar.

    \item François Pailleau, Spring 2023, "Autonomous aerial system for sensor deployment and retrieval in agricultural fields", supervised by Christian Geckeler.

    \item Iris Kong, Spring 2023, "Positioning and Collecting Sensors in Agricultural Fields with Aerial Robots", supervised by Christian Geckeler, part of publication \cite{Geckeler2024a}.
    % best paper award

    \item Sophia Heinrich, Spring 2024, "SnailBot: Robot Locomotion using Biodegradable Adhesives in Forest Canopies", supervised by Christian Geckeler. 
    % TODO part of publication ....
\end{itemize}


\subsection{Semester Projects}
Master students, 3-4 months part time, 8 ECTS
% TODO add 4 students from the Data Science course (ETH Z¨urich), Spring 2022: ”Tree Branch
% Detection: RGB-D Monocular Depth Estimation of Occluded Structure”.
\begin{itemize}
    \item Louis Schalk, Spring 2021, "Development of a robotic system to study the behavior of ants in trees", supervised by Christian Geckeler and Anita Risch

    \item Dario Strübin, Autumn 2021, "Origami-based Miniaturized Aerial Manipulator for Physical Interaction", supervised by Eugenio Cuniato and Christian Geckeler, part of publication \cite{Cuniato2023}

    \item Carlos Augusto Guijo Deza, Autumn 2021, "Magnetic recovery strategy of sensors in a bistable helical origami gripper", supervised by Christian Geckeler, part of publication \cite{Geckeler2022a}

    \item Benito Armas Pizzani, Spring 2022, "Tunable Biodegradable Materials for Robotic Origami", supervised by Christian Geckeler, part of publication \cite{Geckeler2023b}

    \item Gian-Andrea Hehli, Autumn 2022, "Investigating visual-inertial odometry algorithms for forest canopies", supervised by Christian Geckeler.
\end{itemize}

\subsection{Bachelor Thesis}
Bachelor students, 3 months full time, 15 ECTS %3-4 months full time?

\begin{itemize}
  \item Fabian Ospelt, Spring 2021, "Development of an origami manipulator for aerial robots", supervised by Christian Geckeler and Maximilian Brunner, part of publication \cite{Cuniato2023}

  \item Sairaj Rajendra Loke, Autumn 2024, "Visual Estimation of Tree Canopy Rigidity", supervised by Christian Geckeler.
  %TOOD ...?
\end{itemize}


% 2021, 2022 TA Intro Agri. Robo
% 2021, 2022 PhenoRob Workshop, 

% reviewer for: tro, icra, ur, RAL, SoRo

% media coverage: zdv TV gripper, netherlands house  XPrize 
% best paper awards, etc
% exhibited at Robotx x2
% WSL children's future thing, ausflug von wissenschaftliche winthertuhr